---
title: "Week 8 - Homework"
author: "STAT 420, Summer 2020, D. Unger"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes
  pdf_document: default
urlcolor: cyan
---


# Directions

Students are encouraged to work together on homework. However, sharing, copying or providing any part of a homework solution or code is an infraction of the University's rules on Academic Integrity. Any violation will be punished as severely as possible.

- Be sure to remove this section if you use this `.Rmd` file as a template.
- You may leave the questions in your final document.

***

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.alin = "center")
```

## Exercise 1 (Writing Functions)

**(a)** Write a function named `diagnostics` that takes as input the arguments:

- `model`, an object of class `lm()`, that is a model fit via `lm()`
- `pcol`, for controlling point colors in plots, with a default value of `grey`
- `lcol`, for controlling line colors in plots, with a default value of `dodgerblue`
- `alpha`, the significance level of any test that will be performed inside the function, with a default value of `0.05`
- `plotit`, a logical value for controlling display of plots with default value `TRUE`
- `testit`, a logical value for controlling outputting the results of tests with default value `TRUE`

The function should output:

- A list with two elements when `testit` is `TRUE`:
    - `p_val`, the p-value for the Shapiro-Wilk test for assessing normality
    - `decision`, the decision made when performing the Shapiro-Wilk test using the `alpha` value input to the function. "Reject" if the null hypothesis is rejected, otherwise "Fail to Reject."
- Two plots, side-by-side, when `plotit` is `TRUE`:
    - A fitted versus residuals plot that adds a horizontal line at $y = 0$, and labels the $x$-axis "Fitted" and the $y$-axis "Residuals." The points and line should be colored according to the input arguments. Give the plot a title. 
    - A Normal Q-Q plot of the residuals that adds the appropriate line using `qqline()`. The points and line should be colored according to the input arguments. Be sure the plot has a title. 

Consider using this function to help with the remainder of the assignment as well.

``` {r}
diagnostics = function(model, pcol = "grey", lcol = "dodgerblue", alpha = 0.05, plotit = TRUE, testit = TRUE) {
  if (plotit) {
    par (mfrow = c(1,2))
    plot(fitted(model), resid(model), col = pcol, pch = 10, xlab = "Fitted", ylab = "Residuals") 
    abline(h = 0, col = lcol, lwd = 1)
    qqnorm(resid(model), main = "Noraml Q-Q plot of resids", col = pcol)
    qqline(resid(model), col = lcol, lwd = 1)
  }
  
  if (testit) {
    p = shapiro.test(resid(model))$p.value
    deci = ifelse(p < alpha, "Reject", "FTR")
    list(p_val = p, decision = deci)
  }
}
```

  
**(b)** Run the following code.

```{r}
set.seed(40)

data_1 = data.frame(x = runif(n = 30, min = 0, max = 10),
                    y = rep(x = 0, times = 30))
data_1$y = with(data_1, 2 + 1 * x + rexp(n = 30))
fit_1 = lm(y ~ x, data = data_1)

data_2 = data.frame(x = runif(n = 20, min = 0, max = 10),
                    y = rep(x = 0, times = 20))
data_2$y = with(data_2, 5 + 2 * x + rnorm(n = 20))
fit_2 = lm(y ~ x, data = data_2)

data_3 = data.frame(x = runif(n = 40, min = 0, max = 10),
                    y = rep(x = 0, times = 40))
data_3$y = with(data_3, 2 + 1 * x + rnorm(n = 40, sd = x))
fit_3 = lm(y ~ x, data = data_3)
```

```{r, eval = TRUE}
diagnostics(fit_1, plotit = FALSE)$p_val
diagnostics(fit_2, plotit = FALSE)$decision
diagnostics(fit_1, testit = FALSE, pcol = "black", lcol = "black")
diagnostics(fit_2, testit = FALSE, pcol = "grey", lcol = "green")
diagnostics(fit_3)
```

***

## Exercise 2 (Prostate Cancer Data)

For this exercise, we will use the `prostate` data, which can be found in the `faraway` package. After loading the `faraway` package, use `?prostate` to learn about this dataset.

```{r, message = FALSE, warning = FALSE}
library(faraway)
```

**(a)** Fit an additive multiple regression model with `lpsa` as the response and the remaining variables in the `prostate` dataset as predictors. Report the $R^2$ value for this model.

``` {r}
prostate_mlm = lm(lpsa ~ ., data = prostate)
summary(prostate_mlm)$r.squared
```

**(b)** Check the constant variance assumption for this model. Do you feel it has been violated? Justify your answer.
``` {r, warning = FALSE, message = FALSE}
prostate_mlm = lm(lpsa ~ ., data = prostate)
diagnostics(prostate_mlm, testit = FALSE)

library(lmtest)
bptest(prostate_mlm)
```

We can notice from Residuals~Fitted graph, the residual value is largely at 0 value and the residual values spread-out is as expected. We can conclude that the constant variance is not violated.

As for the $bptest$, the p-value is quite large, ascertaining that the constant variance assumption
isn't violated.


**(c)** Check the normality assumption for this model. Do you feel it has been violated? Justify your answer.

``` {r}
prostate_mlm = lm(lpsa ~ ., data = prostate)
shapiro.test(resid(prostate_mlm))$p.value
```
As per the $Shapiro-test$, pvalue is quite high, hence we conclude that "FTR null hypothesis" and normality is not violated.

From the Normal Q-Q Plot, we notice that all the Sample values are around the straight line, to ascertain that the residuals follow a normal distribution.


**(d)** Check for any high leverage observations. Report any observations you determine to have high leverage.
``` {r}
prostate_mlm = lm(lpsa ~ ., data = prostate)
sum(hatvalues(prostate_mlm) > 2 * mean(hatvalues(prostate_mlm)))
prostate[hatvalues(prostate_mlm) > 2 * mean(hatvalues(prostate_mlm)),]
```

There are a total of 5 high leverage observations. 
High leverage observations are calculated based on a beta_hat_value is greater than 2 * mean(beta_hat_values)


**(e)** Check for any influential observations. Report any observations you determine to be influential.
``` {r}
prostate_mlm = lm(lpsa ~ ., data = prostate)

cd = cooks.distance(prostate_mlm)
influ_cd = cd > 4 / length(cd)
sum(influ_cd)
prostate[influ_cd, ]
```

We calculated Influential Observations as observations that have "cook's distance" greater than 4/length(cook's distance entries).
We notice a total of 7 Observations.


**(f)** Refit the additive multiple regression model without any points you identified as influential. Compare the coefficients of this fitted model to the previously fitted model.

``` {r}
refit_prostate_mlm = lm(lpsa ~ ., data = prostate, subset = (cd <= 4/length(cd)))
coef(refit_prostate_mlm)
coef(prostate_mlm)

``` 

Between the refitted model and the original model we notice that intercepts has changed the signs.


**(g)** Create a data frame that stores the observations that were "removed" because they were influential. Use the two models you have fit to make predictions with these observations. Comment on the difference between these two sets of predictions.

``` {r}
influ_data = prostate[influ_cd, ]
predict(prostate_mlm, newdata = influ_data)
predict(refit_prostate_mlm, newdata = influ_data)

predict(refit_prostate_mlm, newdata = influ_data) / predict(prostate_mlm, newdata = influ_data) 
```

There seems to be roughly 10% difference in the predictions between these two models.


***

## Exercise 3 (Why Bother?)

**Why** do we care about violations of assumptions? One key reason is that the distributions of the parameter esimators that we have used are all reliant on these assumptions. When the assumptions are violated, the distributional results are not correct, so our tests are garbage. **Garbage In, Garbage Out!**

Consider the following setup that we will use for the remainder of the exercise. We choose a sample size of 50.

```{r}
n = 50
set.seed(420)
x_1 = runif(n, 0, 5)
x_2 = runif(n, -2, 2)
```

Consider the model,

\[
Y = 4 + 1 x_1 + 0 x_2 + \epsilon.
\]

That is,

- $\beta_0$ = 4
- $\beta_1$ = 1
- $\beta_2$ = 0

We now simulate `y_1` in a manner that does **not** violate any assumptions, which we will verify. In this case $\epsilon \sim N(0, 1).$

```{r}
set.seed(83)
library(lmtest)
y_1 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = 1)
fit_1 = lm(y_1 ~ x_1 + x_2)
bptest(fit_1)
summary(fit_1)
```

Then, we simulate `y_2` in a manner that **does** violate assumptions, which we again verify. In this case $\epsilon \sim N(0, \sigma = |x_2|).$

```{r}
set.seed(83)
y_2 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = abs(x_2))
fit_2 = lm(y_2 ~ x_1 + x_2)
bptest(fit_2)
```

**(a)** Use the following code after changing `birthday` to your birthday.

```{r}
num_sims = 2500
p_val_1 = rep(0, num_sims)
p_val_2 = rep(0, num_sims)
birthday = 19081014
set.seed(birthday)
```

Repeat the above process of generating `y_1` and `y_2` as defined above, and fit models with each as the response `2500` times. Each time, store the p-value for testing,

\[
\beta_2 = 0,
\]

using both models, in the appropriate variables defined above. (You do not need to use a data frame as we have in the past. Although, feel free to modify the code to instead use a data frame.)

``` {r}

for (i in 1: num_sims) {
  y_1 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = 1)
  fit_1 = lm(y_1 ~ x_1 + x_2)
  
  y_2 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = abs(x_2))
  fit_2 = lm(y_2 ~ x_1 + x_2)
  
  p_val_1[i] = summary(fit_1)$coefficients["x_2", "Pr(>|t|)"]
  p_val_2[i] = summary(fit_2)$coefficients["x_2", "Pr(>|t|)"]
}

``` 

**(b)** What proportion of the `p_val_1` values is less than 0.01? Less than 0.05? Less than 0.10? What proportion of the `p_val_2` values is less than 0.01? Less than 0.05? Less than 0.10? Arrange your results in a table. Briefly explain these results.

``` {r}
p1 = c(mean(p_val_1 < 0.01), mean(p_val_1 < 0.05), mean(p_val_1 < 0.10))
p2 = c(mean(p_val_2 < 0.01), mean(p_val_2 < 0.05), mean(p_val_2 < 0.10))
table = rbind(p1, p2)
colnames(table) = c("< 0.01", "< 0.05", "< 0.10")
rownames(table) = c("Model 1", "Model 2")

knitr::kable(table)
```

We notice that pvalues of Model 2 has high proportion whose values are higher than threshold in comparison with Model 1. These higher pvalues creates a possibility that Model 2 could possibily violate Constant Variance assumption and also reject Null Hypothesis.


***

## Exercise 4 (Corrosion Data)

For this exercise, we will use the `corrosion` data, which can be found in the `faraway` package. After loading the `faraway` package, use `?corrosion` to learn about this dataset.

```{r, message = FALSE, warning = FALSE}
library(faraway)
```

**(a)** Fit a simple linear regression with `loss` as the response and `Fe` as the predictor. Plot a scatterplot and add the fitted line. Check the assumptions of this model.

``` {r}
corrosion_lm = lm(loss ~ Fe, data = corrosion)
plot(loss ~ Fe, data = corrosion, col = "blue", pch = 20, cex = 1, main = "Corrosion Loss vs Fe")
abline(corrosion_lm, col = "blue", lwd = 1)

diagnostics(corrosion_lm, testit = FALSE)

bptest(corrosion_lm)
shapiro.test(resid(corrosion_lm))$p.value
```

Constant Variance Assumption

The residual vs Fitted graph shows that the residual value hovers around zero value.
We do notice that the value are distributed as expected.
As for the BPTest results, p-value is shown as high value, hence we can conclude that the Constant Variance assumption is not violated.

Normality Assumption

As for the Q-Q plot, the values seems to fit around the straight line (largely).
Shapiro-test shows a high p-value suggesting that the Normality Assumption isn't violated.

**(b)** Fit higher order polynomial models of degree 2, 3, and 4. For each, plot a fitted versus residuals plot and comment on the constant variance assumption. Based on those plots, which of these three models do you think are acceptable? Use a statistical test(s) to compare the models you just chose. Based on the test, which is preferred? Check the normality assumption of this model. Identify any influential observations of this model.

```{r}
corrosion_2lm = lm(loss ~ Fe + I(Fe ^ 2), data = corrosion)
corrosion_3lm = lm(loss ~ Fe + I(Fe ^ 2) + I(Fe ^ 3), data = corrosion)
corrosion_4lm = lm(loss ~ Fe + I(Fe ^ 2) + I(Fe ^ 3) + I(Fe ^ 4), data = corrosion)

par(mfrow = c(1, 3))
plot(fitted(corrosion_2lm), resid(corrosion_2lm), pch = 20, col = "blue")
abline(h = 0, col = "blue", lwd = 1)

plot(fitted(corrosion_3lm), resid(corrosion_3lm), pch = 20, col = "blue")
abline(h = 0, col = "blue", lwd = 1)

plot(fitted(corrosion_4lm), resid(corrosion_4lm), pch = 20, col = "blue")
abline(h = 0, col = "blue", lwd = 1)

diagnostics(corrosion_2lm, testit = FALSE)
diagnostics(corrosion_3lm, testit = FALSE)
diagnostics(corrosion_4lm, testit = FALSE)

anova(corrosion_2lm, corrosion_3lm)

anova(corrosion_3lm, corrosion_4lm)

pval = c(bptest(corrosion_2lm)$p.value, bptest(corrosion_3lm)$p.value, bptest(corrosion_4lm)$p.value)
st2 = shapiro.test(resid(corrosion_2lm))
st3 = shapiro.test(resid(corrosion_3lm))
st4 = shapiro.test(resid(corrosion_4lm))

sha = c(st2$p.value, st3$p.value, st4$p.value)

ptable = rbind(pval, sha)
colnames(ptable) = c("Model 2", "Model 3", "Model 4")
rownames(ptable) = c("BP P Values", "Shapiro P Values")
knitr::kable(ptable)


```

Based on the Residual vs Fitted graph and Q-Q graphs, we notice that Model 3 - "corrosion_3lm" model seems to be the preferred model. It doesn't seem to violate Constant Variance assumption and the Normality Assumption. This is supported from the result from anova test. We notice a small p-value for Model 3.

As illustrated from the table with p-values from BP Test and Shapiro Test, Model 3 doesn't violate Constant Variance assumption and Normality Assumption.

***

## Exercise 5 (Diamonds)

The data set `diamonds` from the `ggplot2` package contains prices and characteristics of 54,000 diamonds. For this exercise, use `price` as the response variable $y$, and `carat` as the predictor $x$. Use `?diamonds` to learn more.

```{r, message = FALSE, warning = FALSE}
library(ggplot2)
```

**(a)** Fit a linear model with `price` as the response variable $y$, and `carat` as the predictor $x$. Return the summary information of this model.

``` {r}
diamond_lm = lm(price ~ carat, data = diamonds)
summary(diamond_lm)

```

**(b)** Plot a scatterplot of price versus carat and add the line for the fitted model in part **(a)**. Using a fitted versus residuals plot and/or a Q-Q plot, comment on the diagnostics. 

``` {r}
plot(price ~ carat, data = diamonds, col = "blue")
abline(diamond_lm, col = "black", lwd = 3)

plot(fitted(diamond_lm), resid(diamond_lm), col = "blue")
abline(h = 0, col = "black", lwd = 3)

qqnorm(resid(diamond_lm), col = "blue")
qqline(resid(diamond_lm), colo = "black", lwd = 3)
```

As we notice from the Price ~ carat graph, the observations aren't concentrated around the fitted value as we would expect. It is safe to assume that the constant variance assumption is violated.
This can be validated from the Residual vs fitted graph as the observations aren't spread as expected.
As we can notice from the Normal Q-Q plot, we noted that the observations doesn't spread close to the fitted line. It is safe to assume that the Normality assumption is also violated.


**(c)** Seeing as the price stretches over several orders of magnitude, it seems reasonable to try a log transformation of the response. Fit a model with a logged response, plot a scatterplot of log-price versus carat and add the line for the fitted model, then use a fitted versus residuals plot and/or a Q-Q plot to comment on the diagnostics of the model.

```{r}
qplot(price, data = diamonds, bins = 30)

diamond_loglm = lm(log(price) ~ carat, data = diamonds)
plot(log(price) ~ carat, data = diamonds, col = "blue", pch = 20)
abline(diamond_loglm, col = "black", lwd = 3)

plot(fitted(diamond_loglm), resid(diamond_loglm), col = "blue", pch = 20)
abline(h = 0, col = "black", lwd = 2)

qqnorm(resid(diamond_loglm), col = "blue")
qqline(resid(diamond_loglm), col = "black", lwd = 3)

```

We notice from the Residual vs Fitted graph that the Constant Variance Assumption is violated. As the fitted value increase we notice that the values aren't observed around the line
Normality Assumption is also violated as observed from Normal Q-Q Plot.

**(d)** Try adding log transformation of the predictor. Fit a model with a logged response and logged predictor, plot a scatterplot of log-price versus log-carat and add the line for the fitted model, then use a fitted versus residuals plot and/or a Q-Q plot to comment on the diagnostics of the model.

``` {r}
diamond_lllm = lm(log(price) ~ log(carat), data = diamonds)

plot(log(price) ~ log(carat), data = diamonds, col = "blue", pch = 20)
abline(diamond_lllm, col = "black", lwd = 3)

plot(fitted(diamond_lllm), resid(diamond_lllm), col = "blue", pch = 20)
abline(h = 0, col = "black", lwd = 3)

qqnorm(resid(diamond_lllm), col = "blue")
qqline(resid(diamond_lllm), col = "black", lwd = 3)
```

We notice that the observations are plotted reasonable even in loge(price) vs log(carat).
As for the residual vs fitted graph, we notice that observations reasonbly around the fitted line. Not an ideal scenario, but much better than previous model plots. We can safely say that Constant variance assumption isn't violated.

Normal Q-Q plot shows that the observations are around the fitted line and can safely assume that the Normality assumption isn't violated.


**(e)** Use the model from part **(d)** to predict the price (in dollars) of a 3-carat diamond. Construct a 99% prediction interval for the price (in dollars).

``` {r}
exp(predict(diamond_lllm, newdata = data.frame(carat = log(3)), interval = "prediction", level = 0.99))

```
